---
title: "Roberts TS Final Data Preprocessing and Exploration"
author: "Madeleine Roberts"
date: "2024-05-20"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---
## Results:

1.  From plotting the data, we see that we should likely log our data to account for different levels of variance at different time periods.

2.  From the ACF and PCF plots, we see that it is extremely important that our models account for seasonality.

3.  From the Dickey-Fuller and KPSS tests, we see that it that our models should account for differencing in order to make it stationary.

4. We will use tsoutliers to detect and impute outliers at each junction.

5. We see that for all junctions there is a heavy skew to the right.


## Data Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library(readr)
library("Metrics")
library(ggplot2)
library(ggpubr)
library(fpp3)
library(forecast)
options(warn = -1)
library(fable)
library(fabletools)
library(tseries)
```

```{r, include=FALSE}
# Clear working memory
rm(list=ls())

# Set the working directory.
setwd("/Users/madeleineroberts/rWorkingDirectory")
```

```{r}
traffic <- read.csv("traffic.csv")

# Separate into different junctions 
traffic_j1 <- traffic %>% 
  filter(Junction == 1)

traffic_j2 <- traffic %>% 
  filter(Junction == 2)

traffic_j3 <- traffic %>% 
  filter(Junction == 3)

traffic_j4 <- traffic %>% 
  filter(Junction == 4)

# Change time column into a frequency
traffic_j1 <- traffic_j1 %>% 
  mutate(Time = row_number()) 

traffic_j2 <- traffic_j2 %>% 
  mutate(Time = row_number()) 

traffic_j3 <- traffic_j3 %>% 
  mutate(Time = row_number()) 

traffic_j4 <- traffic_j4 %>% 
  mutate(Time = row_number()) 

# Convert dataframe to a time series 
traffic_ts_j1 <- as_tsibble(traffic_j1, index = Time)
traffic_ts_j2 <- as_tsibble(traffic_j2, index = Time)
traffic_ts_j3 <- as_tsibble(traffic_j3, index = Time)
traffic_ts_j4 <- as_tsibble(traffic_j4, index = Time)
```

## Junction 1

**Plot data:**

```{r}
traffic_ts_j1 %>% 
  autoplot(Vehicles) +
  labs(title = "TS of Vehicles at Junction 1")
```

```{r}
traffic_ts_j1 %>% 
  autoplot(log(Vehicles)) +
  labs(title = "TS of Vehicles at Junction 1")
```

Logging the data helps with the heteroskdasticity

**Looking at Autocorrelation Function**

```{r}
ggAcf(traffic_ts_j1$Vehicles)
pacf(traffic_ts_j1$Vehicles)
ggAcf(log(traffic_ts_j1$Vehicles))
pacf(log(traffic_ts_j1$Vehicles))
```

The ACF plot shows significant autocorrelations at multiple lags, which indicates a persistent, long-term correlation structure.
This can suggest that the series is not stationary, as stationary time series typically have autocorrelations that diminish quickly.

The PACF plot shows that the partial autocorrelations drop off quickly after a few lags, which might indicate that the series could be autoregressive (AR) with a finite order.

**Perform the Dickey-Fuller test:**

```{r}
adf_result <- adf.test(log(traffic_ts_j1$Vehicles))
print(adf_result)
```

The Dickey-Fuller test was the first statistical test developed to test the null hypothesis that a unit root is present in an autoregressive model of a given time series, and that the process is thus not stationary.

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely stationary.

**Perform the KPSS Test:**

```{r}
kpss_result <- kpss.test(log(traffic_ts_j1$Vehicles))
print(kpss_result)
```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

ADF Test: Suggests the series is stationary.
KPSS Test: Suggests the series is non-stationary.

**Difference the data to see if we can resolve the stationary issue:**

```{r}
# Differencing the series
traffic_j1_diff <- diff(log(traffic_j1$Vehicles))

# Perform the ADF test on differenced series
adf_result_diff <- adf.test(traffic_j1_diff)
print(adf_result_diff)

# Perform the KPSS test on differenced series
kpss_result_diff <- kpss.test(traffic_j1_diff)
print(kpss_result_diff)
```

Given the p-value of 0.1, which is greater than the common significance level of 0.05, we fail to reject the null hypothesis.
This suggests that the series traffic_j1_diff is likely level stationary.
There is not enough evidence to conclude that the series has a unit root, indicating that differencing the data might have successfully achieved stationarity.

**Detrend the data to see if we can resolve the stationary issue:**

```{r}
# Detrending the series
traffic_j1_detrended <- residuals(lm(log(traffic_j1$Vehicles) ~ time(log(traffic_j1$Vehicles))))

# Perform the ADF test on detrended series
adf_result_detrended <- adf.test(traffic_j1_detrended)
print(adf_result_detrended)

# Perform the KPSS test on detrended series
kpss_result_detrended <- kpss.test(traffic_j1_detrended)
print(kpss_result_detrended)

```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

This suggests that we should difference the data to make it stationary.

**Check for outliers**

```{r}
traffic_ts_j1 <- traffic_ts_j1 %>%
  mutate(Z_score = (Vehicles - mean(Vehicles, na.rm = TRUE)) / sd(Vehicles, na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j1 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
traffic_ts_j1 <- traffic_ts_j1 %>%
  mutate(Z_score = (log(Vehicles) - mean(log(Vehicles), na.rm = TRUE)) / sd(log(Vehicles), na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j1 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
library(tsoutliers)

# Detect outliers using the tsoutliers package
outliers <- tsoutliers(traffic_ts_j1$Vehicles)

print(outliers)
```

**Data Distribution:**

```{r}
# Plot histogram and density plot of original data
ggplot(traffic_ts_j1, aes(x = Vehicles)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.7) +
  geom_density(color = "red") +
  labs(title = "Histogram and Density Plot of Vehicles", x = "Vehicles", y = "Density")
```

**Summary Statistics**
```{r}
# Summary statistics of original data
summary(traffic_ts_j1$Vehicles)
```


## Junction 2

**Plot the data:**

```{r}
traffic_ts_j2 %>% 
  autoplot(Vehicles) +
  labs(title = "TS of Vehicles at Junction 2")
```

```{r}
traffic_ts_j2 %>% 
  autoplot(log(Vehicles)) +
  labs(title = "TS of Vehicles at Junction 2")
```

Logging the data helps with the heteroskdasticity

**Looking at Autocorrelation Function**

```{r}
ggAcf(traffic_ts_j2$Vehicles)
pacf(traffic_ts_j2$Vehicles)
ggAcf(log(traffic_ts_j2$Vehicles))
pacf(log(traffic_ts_j2$Vehicles))
```

The ACF plot suggests that there is a clear seasonality or cyclic pattern in the traffic data.
The periodic peaks could indicate regular traffic patterns, such as daily or weekly cycles.

The ACF plot shows a significant autocorrelation at lag 1, which slowly decays over time, indicating that each data point is correlated with its previous points.

The PACF plot shows a significant spike at lag 1 and then quickly drops off, with minor spikes at higher lags.
This suggests that the immediate past value has a strong influence on the current value, but after accounting for the lag 1 effect, the influence of further lags diminishes

**Perform the Dickey-Fuller test:**

```{r}
adf_result <- adf.test(log(traffic_ts_j2$Vehicles))
print(adf_result)
```

The Dickey-Fuller test was the first statistical test developed to test the null hypothesis that a unit root is present in an autoregressive model of a given time series, and that the process is thus not stationary.

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely stationary.

**Perform the KPSS Test:**

```{r}
kpss_result <- kpss.test(log(traffic_ts_j2$Vehicles))
print(kpss_result)
```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

ADF Test: Suggests the series is stationary.
KPSS Test: Suggests the series is non-stationary.

**Difference the data to see if we can resolve the stationary issue:**

```{r}
# Differencing the series
traffic_j2_diff <- diff(log(traffic_j2$Vehicles))

# Perform the ADF test on differenced series
adf_result_diff <- adf.test(traffic_j2_diff)
print(adf_result_diff)

# Perform the KPSS test on differenced series
kpss_result_diff <- kpss.test(traffic_j2_diff)
print(kpss_result_diff)
```

Given the p-value of 0.1, which is greater than the common significance level of 0.05, we fail to reject the null hypothesis.
This suggests that the series traffic_j2_diff is likely level stationary.
There is not enough evidence to conclude that the series has a unit root, indicating that differencing the data might have successfully achieved stationarity.

**Detrend the data to see if we can resolve the stationary issue:**

```{r}
# Detrending the series
traffic_j2_detrended <- residuals(lm(log(traffic_j2$Vehicles) ~ time(log(traffic_j2$Vehicles))))

# Perform the ADF test on detrended series
adf_result_detrended <- adf.test(traffic_j2_detrended)
print(adf_result_detrended)

# Perform the KPSS test on detrended series
kpss_result_detrended <- kpss.test(traffic_j2_detrended)
print(kpss_result_detrended)

```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

This suggests that we should difference the data to make it stationary.

**Check for outliers**

```{r}
traffic_ts_j2 <- traffic_ts_j2 %>%
  mutate(Z_score = (Vehicles - mean(Vehicles, na.rm = TRUE)) / sd(Vehicles, na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j2 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
traffic_ts_j2 <- traffic_ts_j2 %>%
  mutate(Z_score = (log(Vehicles) - mean(log(Vehicles), na.rm = TRUE)) / sd(log(Vehicles), na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j2 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
library(tsoutliers)

# Detect outliers using the tsoutliers package
outliers <- tsoutliers(traffic_ts_j2$Vehicles)

print(outliers)
```

**Data Distribution:**
```{r}
# Plot histogram and density plot of original data
ggplot(traffic_ts_j2, aes(x = Vehicles)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.7) +
  geom_density(color = "red") +
  labs(title = "Histogram and Density Plot of Vehicles", x = "Vehicles", y = "Density")
```


**Summary Statistics**
```{r}
# Summary statistics of original data
summary(traffic_ts_j2$Vehicles)
```

## Junction 3

**Plot the data:**

```{r}
traffic_ts_j3 %>% 
  autoplot(Vehicles) +
  labs(title = "TS of Vehicles at junction 3")
```

```{r}
traffic_ts_j3 %>% 
  autoplot(log(Vehicles)) +
  labs(title = "TS of Vehicles at junction 3")
```

Logging the data helps with the heteroskdasticity

**Looking at Autocorrelation Function**

```{r}
ggAcf(traffic_ts_j3$Vehicles)
pacf(traffic_ts_j3$Vehicles)
ggAcf(log(traffic_ts_j3$Vehicles))
pacf(log(traffic_ts_j3$Vehicles))
```

The ACF plot shows significant autocorrelations at multiple lags, which indicates a persistent, long-term correlation structure.
This can suggest that the series is not stationary, as stationary time series typically have autocorrelations that diminish quickly.

The PACF plot shows that the partial autocorrelations drop off quickly after a few lags, which might indicate that the series could be autoregressive (AR) with a finite order.

**Perform the Dickey-Fuller test:**

```{r}
adf_result <- adf.test(log(traffic_ts_j3$Vehicles))
print(adf_result)
```

The Dickey-Fuller test was the first statistical test developed to test the null hypothesis that a unit root is present in an autoregressive model of a given time series, and that the process is thus not stationary.

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely stationary.

**Perform the KPSS Test:**

```{r}
kpss_result <- kpss.test(log(traffic_ts_j3$Vehicles))
print(kpss_result)
```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

ADF Test: Suggests the series is stationary.
KPSS Test: Suggests the series is non-stationary.

**Difference the data to see if we can resolve the stationary issue:**

```{r}
# Differencing the series
traffic_j3_diff <- diff(log(traffic_j3$Vehicles))

# Perform the ADF test on differenced series
adf_result_diff <- adf.test(traffic_j3_diff)
print(adf_result_diff)

# Perform the KPSS test on differenced series
kpss_result_diff <- kpss.test(traffic_j3_diff)
print(kpss_result_diff)
```

Given the p-value of 0.1, which is greater than the common significance level of 0.05, we fail to reject the null hypothesis.
This suggests that the series traffic_j3_diff is likely level stationary.
There is not enough evidence to conclude that the series has a unit root, indicating that differencing the data might have successfully achieved stationarity.

**Detrend the data to see if we can resolve the stationary issue:**

```{r}
# Detrending the series
traffic_j3_detrended <- residuals(lm(log(traffic_j3$Vehicles) ~ time(log(traffic_j3$Vehicles))))

# Perform the ADF test on detrended series
adf_result_detrended <- adf.test(traffic_j3_detrended)
print(adf_result_detrended)

# Perform the KPSS test on detrended series
kpss_result_detrended <- kpss.test(traffic_j3_detrended)
print(kpss_result_detrended)

```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

This suggests that we should difference the data to make it stationary.

**Check for outliers**

```{r}
traffic_ts_j3 <- traffic_ts_j3 %>%
  mutate(Z_score = (Vehicles - mean(Vehicles, na.rm = TRUE)) / sd(Vehicles, na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j3 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
traffic_ts_j3 <- traffic_ts_j3 %>%
  mutate(Z_score = (log(Vehicles) - mean(log(Vehicles), na.rm = TRUE)) / sd(log(Vehicles), na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j3 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
library(tsoutliers)

# Detect outliers using the tsoutliers package
outliers <- tsoutliers(traffic_ts_j3$Vehicles)

print(outliers)
```

**Data Distribution:**

```{r}
# Plot histogram and density plot of original data
ggplot(traffic_ts_j3, aes(x = Vehicles)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.7) +
  geom_density(color = "red") +
  labs(title = "Histogram and Density Plot of Vehicles", x = "Vehicles", y = "Density")
```


**Summary Statistics**
```{r}
# Summary statistics of original data
summary(traffic_ts_j3$Vehicles)
```


## Junction 4

**Plot data:**

```{r}
traffic_ts_j4 %>% 
  autoplot(Vehicles) +
  labs(title = "TS of Vehicles at Junction 4")
```

```{r}
traffic_ts_j4 %>% 
  autoplot(log(Vehicles)) +
  labs(title = "TS of Vehicles at Junction 4")
```

Logging the data helps with the heteroskdasticity

**Looking at Autocorrelation Function**

```{r}
ggAcf(traffic_ts_j4$Vehicles)
pacf(traffic_ts_j4$Vehicles)
ggAcf(log(traffic_ts_j4$Vehicles))
pacf(log(traffic_ts_j4$Vehicles))
```

The ACF plot shows significant autocorrelations at multiple lags, which indicates a persistent, long-term correlation structure.
This can suggest that the series is not stationary, as stationary time series typically have autocorrelations that diminish quickly.

The PACF plot shows that the partial autocorrelations drop off quickly after a few lags, which might indicate that the series could be autoregressive (AR) with a finite order.

**Perform the Dickey-Fuller test:**

```{r}
adf_result <- adf.test(log(traffic_ts_j4$Vehicles))
print(adf_result)
```

The Dickey-Fuller test was the first statistical test developed to test the null hypothesis that a unit root is present in an autoregressive model of a given time series, and that the process is thus not stationary.

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely stationary.

**Perform the KPSS Test:**

```{r}
kpss_result <- kpss.test(log(traffic_ts_j4$Vehicles))
print(kpss_result)
```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

ADF Test: Suggests the series is stationary.
KPSS Test: Suggests the series is non-stationary.

**Difference the data to see if we can resolve the stationary issue:**

```{r}
# Differencing the series
traffic_j4_diff <- diff(log(traffic_j4$Vehicles))

# Perform the ADF test on differenced series
adf_result_diff <- adf.test(traffic_j4_diff)
print(adf_result_diff)

# Perform the KPSS test on differenced series
kpss_result_diff <- kpss.test(traffic_j4_diff)
print(kpss_result_diff)
```

Given the p-value of 0.1, which is greater than the common significance level of 0.05, we fail to reject the null hypothesis.
This suggests that the series traffic_j4_diff is likely level stationary.
There is not enough evidence to conclude that the series has a unit root, indicating that differencing the data might have successfully achieved stationarity.

**Detrend the data to see if we can resolve the stationary issue:**

```{r}
# Detrending the series
traffic_j4_detrended <- residuals(lm(log(traffic_j4$Vehicles) ~ time(log(traffic_j4$Vehicles))))

# Perform the ADF test on detrended series
adf_result_detrended <- adf.test(traffic_j4_detrended)
print(adf_result_detrended)

# Perform the KPSS test on detrended series
kpss_result_detrended <- kpss.test(traffic_j4_detrended)
print(kpss_result_detrended)

```

P-value = 0.01: Strong evidence against the null hypothesis.
Decision: Reject the null hypothesis.
Conclusion: The time series is likely non-stationary, suggesting the presence of a unit root.

This suggests that we should difference the data to make it stationary.

**Check for outliers**

```{r}
traffic_ts_j4 <- traffic_ts_j4 %>%
  mutate(Z_score = (Vehicles - mean(Vehicles, na.rm = TRUE)) / sd(Vehicles, na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j4 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
traffic_ts_j4 <- traffic_ts_j4 %>%
  mutate(Z_score = (log(Vehicles) - mean(log(Vehicles), na.rm = TRUE)) / sd(log(Vehicles), na.rm = TRUE))

# Identify outliers
outliers <- traffic_ts_j4 %>%
  filter(abs(Z_score) > 3)

count(outliers)
```

```{r}
library(tsoutliers)

# Detect outliers using the tsoutliers package
outliers <- tsoutliers(traffic_ts_j4$Vehicles)

print(outliers)
```

**Data Distribution:**

```{r}
# Plot histogram and density plot of original data
ggplot(traffic_ts_j4, aes(x = Vehicles)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.7) +
  geom_density(color = "red") +
  labs(title = "Histogram and Density Plot of Vehicles", x = "Vehicles", y = "Density")
```


**Summary Statistics**
```{r}
# Summary statistics of original data
summary(traffic_ts_j4$Vehicles)
```


